---
title: "Basic R proficiency"
author: "Zirong Li"
date: "2026-09-23"
output: html_document
---

#Load libraries

```{r}
library(tidyverse)

```

#Import data

```{r}
mobdata <- read.csv("https://raw.githubusercontent.com/wellsdata/CompText_Jour/main/data/mob_analysis_data.csv")
```

#Describe the number of rows and columns in the dataset

```{r}
glimpse(mobdata)

#Number of rows is 3,836
#Columns are in the table
```

#Create a table that displays a count of the top 15 newspaper entries

```{r}
mobdata %>%
  count(Newspaper) %>%
  slice_max(order_by = n, n = 15)

#There is a tie at the 15th position. If we want exactly 15 entries:
mobdata %>%
  count(Newspaper) %>%
  slice_max(order_by = n, n = 15, with_ties = F)
```

#Create a table that counts the different entries in the "keyword" column

```{r}
mobdata %>%
  count(keyword) %>%
  arrange(desc(n))
```

#Create a table that counts all specific articles by year

```{r}
mobdata %>%
  group_by(Year) %>%
  summarise(article_count = n_distinct(docname)) %>%
  arrange(desc(article_count))
#I use "docname" variable to count the unique articles within each year
```

#Create a simple column ggplot chart that shows the total entries by year

```{r}
mobdata %>%
  count(Year) %>%
  arrange(desc(n)) %>%
  ggplot(aes(x = Year, y = n,
             fill = n)) +
  geom_col(position = "dodge") +
  theme(legend.position = "none") +
  labs(
    title = "Total Newspaper Entries by Year",
    y = "Number of Entries",
    x = "Year"
  )
```

**Memo of findings** The dataset contains 3,836 entries drawn from a range of newspapers and years. These entries are not evenly distributed across years or newspapers. The most frequently represented newspaper is "the Birmingham age-herald (Birmingham, Ala...)" with 118 entries, followed by "the daily worker (Chicago, Ill.)" with 64 entries and "the Topeka state journal (Topeka, Kansas)" with 63 entries.

The keyword frequencies are highly concentrated around a few related concepts. Variants of "mob" (also including "mobs", "mob's", and "mob-") account for most of the entries, while another common group includes "masked", "masked men", and "masks". These categories appear to represent similar concepts but are recorded as separate keyword values, so for a more accurate representation of keyword frequencies, it would be better to combine similar concepts during data cleaning. Now it is clear that two keyword groups, "mob" and "masked" are much more prominent in the collection than others.

Looking at the data over time also reveals variation in coverage. 1903 has the highest number of unique articles, with 120 articles. Looking at the whole distribution, although the dataset covers more than a century, the observations are concentrated around the late nineteenth and early twentieth centuries. Entries increase substantially around the 1890s, reach their highest levels in the early 1900s, and generally decline in later decades. The uneven distribution raises a question about whether the peak reflects a genuine increase in newspaper coverage of the topic of "mob news" or differences in the availability and collection of historical newspaper records.
