---
title: "Metadata"
author: "Wells"
date: "2026-09-15"
output: html_document
---

Class exercise to import your project metadata from the spreadsheet into R 

--Create a folder called "class_project" in your code folder 

--Save this file to the "class_project" folder 

--Move your Nexis spreadsheet to the folder 

--Create a project file on R Studio: File \| New Project \| \*\*\* select Existing folder \| "class_project" 

--After project.Rproj launches, then launch this file, metadata.Rmd, and complete the tutorial

**By doing this, you avoid the ring of hell where R can't find your data**

#load tidyverse, janitor, rio

```{r}

library(tidyverse)
library(janitor)
library(rio)
#install.packages("readxl")
library(readxl)

```

#Import your data

```{r}

data <- read_excel("AI_and_Healthsource.XLSX") |> 
  clean_names()
View(data)

```

# Clean names

```{r}
data <- data |> 
  clean_names()
names(data)
```

# check data types
```{r}
glimpse(data)
```

#clean up the data, as there are many empty and useless columns
```{r}
meta100 <- data |>
  transmute(
    id          = sprintf("A%03d", row_number()),
    title       = str_squish(title),
    date        = as_date(published_date),
    publication = str_squish(publication_4),
    pub_type    = str_squish(publication_type_5),
    word_count  = as.integer(word_count),
    section     = str_squish(section),
    byline      = str_squish(byline),
    company_tag = str_squish(company),
    hlead       = str_squish(hlead)
  )

head(meta100)

```

#save the new metadata excel
```{r}
write_xlsx(meta100, "pilot100_metadata.xlsx")
```


Next Steps: Count years, and create a ggplot

# Transform date into years. See tutorial
```{r}

meta100 <- meta100 |> 
  mutate(year = year(date))

```

#publication
```{r}

meta100 |> 
  count(publication) |>
  arrange(desc(n))

```

#filter on length
```{r}

meta100 |> 
  mutate(length = as.numeric(word_count)) |> 
  filter(length > 800)

```

# Create a ggplot
```{r}

meta100 |> 
  count(year) |> 
   ggplot(aes(x = year, y = n,
             fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  scale_x_continuous(breaks=c(2022:2026)) + #fixes wonky x axis
  labs(title = "Years for News Discourse about AI as Health Source Research",
       subtitle = "Limited NexisUni Run",
       caption = "Graphic by Rob Wells, 9-15-2026",
       y="Number of Articles",
       x="")

#I extracted evenly for each year, so group by year doesn't work for me

```

#Count publications

```{r}
meta100 |> 
  count(publication) |>
  arrange(desc(n))
```


#Visualize publications

```{r}
meta100 |> 
  count(publication) |> 
   ggplot(aes(x = n, y = reorder(publication, n), #change the direction of bars and set the order
             fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  labs(title = "Publications for News Discourse about AI as Health Source Research",
       subtitle = "Limited NexisUni Run",
       caption = "Graphic by Rob Wells, 9-15-2026",
       y="Number of Articles",
       x="Publications")

```



# NOTES ON FIXING COMPLEX Date COLUMN WITH HOURS AND MINUTES

```{r}
women <- women |> 
  mutate(date = as.Date(strptime(published_date, format = "%B %d, %Y %A")))
```